Self Referencing Sequences

### 2014-11-06 draft.tex
finished modifying equivalence section

The length of the block that $a_{m+j-1}$ is a part of is $a_{m+n}$, for some $0 \leq n < j-1$. Because the two sequences are identical for $0 \leq k<j$ and by our assumption that the read distance is the same for $a_{m}$ and $b_{l}$, the read distance for $b_{l+n}$ equals the read distance for $a_{m+n}$. This combined with the fact that $a_{m+n} = b_{l+n}$ shows that the block including $a_{m+j-1}$ must be of the same length as the block including $b_{l+j-1}$.

Since the two sequences are the same for $0 \leq k<j$ the number of elements in the block containing $a_{m+j-1}$, with $k<j$, and the number of elements in the block containing $b_{l+j-1}$, with $k<j$, must be the same. 
Because $a_{m+j}$ is part of the same block as the element preceding it, this means that $b_{l+j}$ must be a part of the same block as the element proceeding it as well. 
This contradicts the fact that exactly one of the two elements is not in the same block as the element preceding it and so the two sequences must be equivalent.

@@

This result also allows us to make a general statement for equivalence of sequences that begin with a $1$.
\begin{corollary} If $C = (1 c_1 c_2 \cdots c_n)$, then $s(C, 1)$ and $s(C, c_1)$ are equivalent.
\end{corollary}
\begin{proof}
Let $\{a_i\} = s(C, 1)$ and $\{b_i\} = s(C, c_1)$. The read distance of sequence $\{a_i\}$ for $i=2$ is 0, and the read distance of sequence $\{b_i\}$ for $i = 1$ is 0. In addition, $a_2 = b_1 = c_1$, so by Theorem ~\ref{equivsuff} the two sequences are equivalent.
\end{proof}

We can also demonstrate a necessary condition for equivalence. 
\begin{theorem}
\label{thm:readfirst}
If two sequences are equivalent then the first element of the shared sequence must have the same read distance in both sequences.
\end{theorem}

@@

Since the two read distances are different we know there are two separate blocks of length $c_1$ in the common sequence, one which corresponds to $a_{m+1}$ and another which corresponds to $b_{l+1}$. Note that the block corresponding to $b_{l+1}$ is to the right of the block corresponding to $a_{m+1}$ because the read distance for $b_{l+1}$ is larger. Now because $\{a_i\}$ is self referential there exists some index $n$ such that $a_{m+n+1}$ is the element of $\{a_i\}$ which corresponds to the second block of length $c_1$.

We claim that
\begin{equation*}
c_i = b_{l-n+i} \quad 1 \leq i \leq n
\end{equation*}
That is the subsequence $c_1, c_2, \ldots, c_{n}$ appears in $\{b_i\}$ such that the last term is next to the beginning of the common sequence. This is true because the length of the blocks of $\{c_i\}$ which occur before the second block of length $c_1$ must appear as terms of $\{b_i\}$ and must end next to the number which refers to the second block of length $c_1$.

Thus our sequence $\{b_i\}$ is  
\begin{equation*}
\cdots c_1 c_2 \cdots c_{n} c_1 c_2 \cdots c_n \cdots
\end{equation*}
where the beginning of the common sequence is the second $c_1$. We now claim that the common sequence is just the subsequence
\begin{equation*}
c_1 c_2 \cdots c_n,
\end{equation*}
repeated an infinite number of times.

@@

\begin{corollary}
If two self-referencing sequences are generated over the same cycle  and are equivalent then the read distance of their shared subsequence is the same everywhere.
\end{corollary}

@@

It is unclear to us whether all sequences generated by the same cycle are equivalent. This is certainly the case for $C=(123)$ and any two element cycles but we have been unable to generalize this to other three element or larger cycles. A key ingredient in establishing a counterexample would be more stringent necessary conditions for equivalence but it is unclear what these conditions should be or how to prove they are necessary. In order to establish that all sequences generated by the same cycle are in fact equivalent we would like to establish weaker or more generalizable sufficient conditions for equivalence, but based on our proof of Theorem ~\ref{equivsuff} it does not seem like weaker conditions will actually be sufficient.

>>>>

The length of the block that $a_{m+j-1}$ is a part of is $a_{m+n}$, for some $0 \leq n < j-1$. Because the two sequences are identical for $0 \leq k<j$ and by our assumption that the read distance is the same for $a_{m}$ and $b_{l}$, the read distance for $b_{l+n}$ equals the read distance for $a_{m+n}$. This combined with the fact that $a_{m+n} = b_{l+n}$ shows that the block containing $a_{m+j-1}$ must be of the same length as the block containing $b_{l+j-1}$.

Since the two sequences are the same for $0 \leq k<j$, the number of elements in the block containing $a_{m+j-1}$, with $k<j$, and the number of elements in the block containing $b_{l+j-1}$, with $k<j$, must be the same. Because $a_{m+j}$ is part of the same block as the element preceding it, this means that $b_{l+j}$ must be a part of the same block as the element proceeding it as well. 
This contradicts the fact that exactly one of the two elements is not in the same block as the element preceding it and so the two sequences must be equivalent.

@@

This result also allows us to make a general statement regarding equivalence of sequences that begin with a $1$.
\begin{corollary} If $C = (1 c_1 c_2 \cdots c_n)$, then $s(C, 1)$ and $s(C, c_1)$ are equivalent.
\end{corollary}
\begin{proof}
Let $\{a_i\} = s(C, 1)$ and $\{b_i\} = s(C, c_1)$. The read distance of sequence $\{a_i\}$ for $i=2$ is 0, and the read distance of sequence $\{b_i\}$ for $i = 1$ is 0. In addition, $a_2 = b_1 = c_1$, so by Theorem ~\ref{equivsuff} the two sequences are equivalent.
\end{proof}
Note that one can check this corollary easily by hand, but our formalism provides a way to mathematically verify this intuition.

We can also demonstrate a necessary condition for equivalence. 
\begin{theorem}
\label{thm:readfirst}
If two sequences are equivalent, then the first element of the shared sequence must have the same read distance in both sequences.
\end{theorem}

@@

Since the two read distances are different, we know there are two separate blocks of length $c_1$ in the common sequence, one which corresponds to $a_{m+1}$ and another which corresponds to $b_{l+1}$. Note that the block corresponding to $b_{l+1}$ is to the right of the block corresponding to $a_{m+1}$ because the read distance for $b_{l+1}$ is larger. Now because $\{a_i\}$ is self-referential, there exists some index $n$ such that $a_{m+n+1}$ is the element of $\{a_i\}$ which corresponds to the second block of length $c_1$.

We claim that
\begin{equation*}
c_i = b_{l-n+i} \quad 1 \leq i \leq n.
\end{equation*}
That is, the subsequence $c_1, c_2, \ldots, c_{n}$ appears in $\{b_i\}$ such that the last term is next to the beginning of the common sequence. This is true because the length of the blocks of $\{c_i\}$ which occur before the second block of length $c_1$ must appear as terms of $\{b_i\}$ and must end next to the number which refers to the second block of length $c_1$.

Thus our sequence $\{b_i\}$ is  
\begin{equation*}
\cdots c_1 c_2 \cdots c_{n} c_1 c_2 \cdots c_n \cdots \ ,
\end{equation*}
where the beginning of the common sequence is the second $c_1$. We now claim that the common sequence is just the subsequence
\begin{equation*}
c_1 c_2 \cdots c_n,
\end{equation*}
repeated an infinite number of times.

@@

\begin{corollary}
If two self-referencing sequences are generated over the same cycle and are equivalent, then the read distance of their shared subsequence is the same everywhere.
\end{corollary}

@@

It is unclear whether all sequences generated by the same cycle are equivalent. This is certainly the case for $C=(123)$ and any two element cycles, but we have been unable to generalize this to other three element or larger cycles. A key ingredient in establishing a counterexample would be more stringent necessary conditions for equivalence, but it is unclear what these conditions should be or how to prove they are necessary. If we want to establish that all sequences generated by the same cycle are in fact equivalent, we would need weaker or more generalizable sufficient conditions for equivalence.
